### 2024-02-17 latex/draft 1/ProtestsDraft1.tex
started latex doc

\documentclass{article}

% Packages for statistical analysis
\usepackage{amsmath} % For mathematical symbols and equations
\usepackage{amssymb} % For additional mathematical symbols
\usepackage{graphicx} % For including figures
\usepackage{caption} % For customizing captions
\usepackage{subcaption} % For subfigures
\usepackage{booktabs} % For professional-looking tables
\usepackage{siunitx} % For typesetting units and quantities
\usepackage{natbib} % For bibliography management

% Title and author information
\title{Statistical Analysis Paper}
\author{Your Name}
\date{\today}

\begin{document}

\maketitle

% Abstract
\begin{abstract}
This is the abstract of your paper.
\end{abstract}

% Introduction
\section{Introduction}
This is the introduction section of your paper.

% Methodology
\section{Methodology}
This is the methodology section of your paper.

% Results
\section{Results}
This is the results section of your paper.

% Discussion
\section{Discussion}
This is the discussion section of your paper.

% Conclusion
\section{Conclusion}
This is the conclusion section of your paper.

% References
\bibliographystyle{plainnat}
\bibliography{references}

\end{document}