### 2024-04-02 latex/draft/draft.tex
gather data

\documentclass[12pt]{article}

% Packages
\usepackage{amsmath} % For mathematical symbols and equations
\usepackage{graphicx} % For including figures
\usepackage{hyperref} % For hyperlinks
\usepackage{listings} % Required for inserting code snippets
\usepackage{xcolor} % Required for custom colors
\usepackage[margin=1in]{geometry}
\usepackage{titling}
\usepackage{setspace}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{enumitem}

\lstset{
    language=R, % Set the language for code snippets to R
    basicstyle=\ttfamily\small, % Set the font and size for code snippets
    keywordstyle=\color{blue}, % Set the color for keywords
    commentstyle=\color{green!60!black}, % Set the color for comments
    stringstyle=\color{red}, % Set the color for strings
    showstringspaces=false, % Don't show spaces in strings
    breaklines=true, % Allow code to break across lines
    frame=single, % Add a frame around code snippets
    numbers=left, % Add line numbers to code snippets
    numberstyle=\tiny\color{gray}, % Set the style for line numbers
    captionpos=b % Position captions at the bottom of code snippets
}

% Title and author
\title{Title}
\author{Leo Kraushaar \& Ali Hay \\ STAT 413}
\date{\today}

\begin{document}
\onehalfspacing

\begin{titlepage}
    \newcommand{\HRule}{\rule{\linewidth}{0.5mm}}
    \center

    \HRule \\[1cm]
    { \huge \bfseries \thetitle}\\[0.4cm] 
    \HRule \\[1.8cm]

    \Large \thedate\\[5cm]

    \begin{minipage}{1\textwidth}
    \vspace{10cm}
    \begin{flushleft} \large
    \theauthor
    \end{flushleft}

    \end{minipage}

\end{titlepage}

\newpage
\section{Introduction}

\newpage
\section{Model Building}
\subsection{Fitting Process}
\subsection{Model Overview}

\newpage
\section{Bootstrapping for Feature Significance}
\subsection{Resampling Bootstrap}
\subsubsection*{Procedure}
\subsubsection*{Results}

\subsection{Parametric Bootstrap}
\subsubsection*{Procedure}
\subsubsection*{Results}

\subsection{Smooth Bootstrap}
\subsubsection*{Procedure}
\subsubsection*{Results}

\subsection{Method Comparison}

\newpage
\section{Monte Carlo Estimation}

\newpage
\section{Conclusion}

\end{document}