### 2024-04-03 latex/draft/draft.tex
new results

\documentclass[12pt]{article}

% Packages
\usepackage{amsmath} % For mathematical symbols and equations
\usepackage{graphicx} % For including figures
\usepackage{hyperref} % For hyperlinks
\usepackage{listings} % Required for inserting code snippets
\usepackage{xcolor} % Required for custom colors
\usepackage[margin=1in]{geometry}
\usepackage{titling}
\usepackage{setspace}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{enumitem}

\lstset{
    language=R, % Set the language for code snippets to R
    basicstyle=\ttfamily\small, % Set the font and size for code snippets
    keywordstyle=\color{blue}, % Set the color for keywords
    commentstyle=\color{green!60!black}, % Set the color for comments
    stringstyle=\color{red}, % Set the color for strings
    showstringspaces=false, % Don't show spaces in strings
    breaklines=true, % Allow code to break across lines
    frame=single, % Add a frame around code snippets
    numbers=left, % Add line numbers to code snippets
    numberstyle=\tiny\color{gray}, % Set the style for line numbers
    captionpos=b % Position captions at the bottom of code snippets
}

% Title and author
\title{Title}
\author{Leo Kraushaar \& Ali Hay \\ STAT 413}
\date{\today}

\begin{document}
\onehalfspacing

\begin{titlepage}
    \newcommand{\HRule}{\rule{\linewidth}{0.5mm}}
    \center

    \HRule \\[1cm]
    { \huge \bfseries \thetitle}\\[0.4cm] 
    \HRule \\[1.8cm]

    \Large \thedate\\[5cm]

    \begin{minipage}{1\textwidth}
    \vspace{10cm}
    \begin{flushleft} \large
    \theauthor
    \end{flushleft}

    \end{minipage}
\end{titlepage}

\newpage
\section{Introduction}

\newpage
\section{Model Building}
\subsection{Fitting Process}
\subsection{Model Overview}

\newpage
\section{Bootstrapping for Feature Significance}
\subsection{Resampling Bootstrap}
\subsubsection*{Procedure}
Using $B = 10000$, an algorithm was written to perform a resampling bootstrap on the dataset. On each iteration, a sample of size $n$ (where $n$ was the number of rows in the dataset) was taken with replacement. Using this sample, a new model was fit. The parameter estimates were recorded.

\vspace{1cm}
\begin{lstlisting}[language=R]
resampBoot <- function(df, B) {
    # Get sample size
    n <- nrow(df)
    # Initialize empty dataframe
    params <- c()
    # Initialize progress bar
    bar <- txtProgressBar(min=0, max=B, style=1)
    # Perform B iterations
    for (b in 1:B) {
        # Select a sample of size n
        indices <- sample(1:n, replace = TRUE)
        samp <- df[indices, ]
        # Fit the model with the sample
        boot_model <- glm.nb(protests ~., data=samp, init.theta = 10)
        boot_params <- coef(boot_model)
        params <- rbind(params, boot_params)
        setTxtProgressBar(bar, b)
    }
    close(bar)
    return(params)
}
\end{lstlisting}

\subsubsection*{Results}

\begin{table}[hb!]
    \centering
    \begin{tabular}{|c|c|c|c|c|c|c|}
        \hline
     & \textbf{mean} & \textbf{sd} & \textbf{2.5\%} & \textbf{50\%} & \textbf{97.5\%} & \textbf{contains 0} \\ 
        \hline
    intercept & 3.62 & 0.49 & 2.64 & 3.63 & 4.57 & FALSE \\ 
      Aug & -0.71 & 0.16 & -1.02 & -0.71 & -0.40 & FALSE \\ 
      Dec & -0.64 & 0.36 & -1.07 & -0.62 & -0.29 & FALSE \\ 
      Feb & 0.04 & 0.16 & -0.26 & 0.04 & 0.36 & TRUE \\ 
      Jan & -0.34 & 0.16 & -0.64 & -0.34 & -0.02 & FALSE \\ 
      Jul & -0.56 & 0.16 & -0.87 & -0.56 & -0.23 & FALSE \\ 
      Jun & -0.35 & 0.13 & -0.61 & -0.35 & -0.09 & FALSE \\ 
      Mar & -0.11 & 0.15 & -0.39 & -0.11 & 0.18 & TRUE \\ 
      May & -0.03 & 0.14 & -0.30 & -0.03 & 0.25 & TRUE \\ 
      Nov & -0.08 & 0.14 & -0.37 & -0.08 & 0.20 & TRUE \\ 
      Oct & 0.04 & 0.14 & -0.23 & 0.03 & 0.32 & TRUE \\ 
      Sep & 0.10 & 0.13 & -0.15 & 0.10 & 0.36 & TRUE \\ 
      BC & 0.82 & 0.13 & 0.55 & 0.82 & 1.08 & FALSE \\ 
      M & -1.87 & 0.87 & -3.56 & -1.89 & -0.11 & FALSE \\ 
      NB & -2.58 & 0.99 & -4.49 & -2.59 & -0.57 & FALSE \\ 
      NL & -3.01 & 1.05 & -5.05 & -3.03 & -0.86 & FALSE \\ 
      NT & -5.37 & 1.33 & -7.81 & -5.37 & -2.90 & FALSE \\ 
      NS & -2.36 & 0.94 & -4.18 & -2.37 & -0.44 & FALSE \\ 
      N & -4.95 & 1.20 & -7.28 & -4.97 & -2.53 & FALSE \\ 
      O & 5.49 & 2.23 & 0.96 & 5.52 & 9.81 & FALSE \\ 
      PEI & -4.07 & 1.16 & -6.30 & -4.07 & -1.72 & FALSE \\ 
      Q & 2.16 & 0.83 & 0.49 & 2.17 & 3.79 & FALSE \\ 
      S & -2.54 & 0.90 & -4.29 & -2.56 & -0.73 & FALSE \\ 
      Y & -4.08 & 1.18 & -6.39 & -4.10 & -1.70 & FALSE \\ 
      retail & -1.80 & 0.98 & -3.70 & -1.81 & 0.18 & TRUE \\ 
      \hline
    \end{tabular}
\end{table}

\newpage
\subsection{Parametric Bootstrap}
\subsubsection*{Procedure}
\subsubsection*{Results}

\subsection{Smooth Bootstrap}
\subsubsection*{Procedure}
\subsubsection*{Results}

\subsection{Method Comparison}

\newpage
\section{Monte Carlo Estimation}

\newpage
\section{Conclusion}

\end{document}