### 2024-04-03 latex/draft/draft.tex
added results to tex file

\documentclass[12pt]{article}

% Packages
\usepackage{amsmath} % For mathematical symbols and equations
\usepackage{graphicx} % For including figures
\usepackage{hyperref} % For hyperlinks
\usepackage{listings} % Required for inserting code snippets
\usepackage{xcolor} % Required for custom colors
\usepackage[margin=1in]{geometry}
\usepackage{titling}
\usepackage{setspace}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{enumitem}

\lstset{
    language=R, % Set the language for code snippets to R
    basicstyle=\ttfamily\small, % Set the font and size for code snippets
    commentstyle=\color{green!60!black}, % Set the color for comments
    stringstyle=\color{orange}, % Set the color for strings
    showstringspaces=false, % Don't show spaces in strings
    breaklines=true, % Allow code to break across lines
    frame=single, % Add a frame around code snippets
    numbers=left, % Add line numbers to code snippets
    numberstyle=\tiny\color{gray}, % Set the style for line numbers
    captionpos=b % Position captions at the bottom of code snippets
}

% Title and author
\title{Title}
\author{Leo Kraushaar \& Ali Hay \\ STAT 413}
\date{\today}

\begin{document}
\onehalfspacing

\begin{titlepage}
    \newcommand{\HRule}{\rule{\linewidth}{0.5mm}}
    \center

    \HRule \\[1cm]
    { \huge \bfseries \thetitle}\\[0.4cm] 
    \HRule \\[1.8cm]

    \Large \thedate\\[5cm]

    \begin{minipage}{1\textwidth}
    \vspace{10cm}
    \begin{flushleft} \large
    \theauthor
    \end{flushleft}

    \end{minipage}
\end{titlepage}

\newpage
\section{Introduction}

\newpage
\section{Model Building}
\subsection{Fitting Process}
\subsection{Model Overview}

\newpage
\section{Bootstrapping for Feature Significance}
\subsection{Resampling Bootstrap}
\subsubsection*{Procedure}
Using $B = 10000$, an algorithm was written to perform a resampling bootstrap on the dataset. On each iteration, a sample of size $n$ (where $n$ was the number of rows in the dataset) was taken with replacement. Using this sample, a new model was fit. The parameter estimates were recorded. Relevant code is shown below.

\vspace{1cm}
\begin{lstlisting}[language=R]
resampBoot <- function(df, B) {
    # Get sample size
    n <- nrow(df)
    # Initialize empty dataframe
    params <- c()
    # Initialize progress bar
    bar <- txtProgressBar(min=0, max=B, style=1)
    # Perform B iterations
    for (b in 1:B) {
        # Select a sample of size n
        indices <- sample(1:n, replace = TRUE)
        samp <- df[indices, ]
        # Fit the model with the sample
        boot_model <- glm.nb(protests ~., data=samp, init.theta = 10)
        boot_params <- coef(boot_model)
        params <- rbind(params, boot_params)
        setTxtProgressBar(bar, b)
    }
    close(bar)
    return(params)
}
\end{lstlisting}

\subsubsection*{Results}
\begin{table}[hbt!]
    \centering
    \begin{tabular}{|l|c|c|c|c|c|c|}
      \hline
     & mean & sd & 2.5\% & 50\% & 97.5\% & sig \\ 
      \hline
    intercept & 3.66 & 0.53 & 2.70 & 3.64 & 4.77 & TRUE \\ 
      Spr & -0.06 & 0.08 & -0.23 & -0.06 & 0.09 & FALSE \\ 
      Sum & -0.55 & 0.08 & -0.72 & -0.55 & -0.38 & TRUE \\ 
      Win & -0.24 & 0.10 & -0.42 & -0.23 & -0.05 & TRUE \\ 
      BC & 0.82 & 0.14 & 0.55 & 0.82 & 1.09 & TRUE \\ 
      M & -1.89 & 0.93 & -3.85 & -1.85 & -0.18 & TRUE \\ 
      NB & -2.60 & 1.06 & -4.84 & -2.55 & -0.66 & TRUE \\ 
      NL & -3.04 & 1.13 & -5.43 & -2.98 & -0.99 & TRUE \\ 
      NT & -5.40 & 1.37 & -8.20 & -5.34 & -2.94 & TRUE \\ 
      NS & -2.38 & 1.00 & -4.53 & -2.34 & -0.54 & TRUE \\ 
      N & -4.98 & 1.28 & -7.68 & -4.92 & -2.61 & TRUE \\ 
      O & 5.54 & 2.37 & 1.18 & 5.40 & 10.58 & TRUE \\ 
      PEI & -4.10 & 1.23 & -6.73 & -4.04 & -1.82 & TRUE \\ 
      Q & 2.17 & 0.88 & 0.55 & 2.12 & 4.04 & TRUE \\ 
      S & -2.56 & 0.96 & -4.57 & -2.52 & -0.77 & TRUE \\ 
      Y & -4.11 & 1.26 & -6.77 & -4.06 & -1.81 & TRUE \\ 
      retail & -1.83 & 1.04 & -4.05 & -1.77 & 0.08 & FALSE \\ 
       \hline
    \end{tabular}
\end{table}
\newpage
\subsection{Parametric Bootstrap}
\subsubsection*{Procedure}
Using $B = 10000$, an algorithm was written to perform a parametric bootstrap on the dataset.

Using the estimated dispersion parameter $\theta \approx 8.36$, each iteration sampled a random vector from a negative binomial distribution. The distribution had dispersion parameter $\theta$ and mean $\hat y$, where $\hat y$ was the predicted mean value for the corresponding input values.

Using these new estimates, a model was fit on each iteration and the parameter estimates were recorded. Relevant code is shown below.

\begin{lstlisting}
conditionalNegBinom <- function(theta, mu) {
    nb_sample <- rnbinom(size=theta, mu=mu, n=1)
    return(nb_sample)
}

paramBoot <- function(B, X, yhat, theta, func) {

    # Initialize empty vector
    params <- c()
    # Iterate B times
    for (b in 1:B) {
        # Simulate NB given means
        sim_y <- sapply(yhat, function(y) func(theta, y))
        # Add to the dataframe
        sim_data <- cbind(X, protests=sim_y)
        # Fit the model to the simulated data
        sim_model <- glm.nb(protests ~., data=sim_data, init.theta = theta)
        # Access the coefficients and store
        parameters <- coef(sim_model)
        params <- rbind(params, parameters)
    }
    return(params)
}
\end{lstlisting}
\subsubsection*{Results}
\begin{table}[hbt!]
    \centering
    \begin{tabular}{|l|c|c|c|c|c|c|}
      \hline
     & mean & sd & 2.5\% & 50\% & 97.5\% & sig \\ 
      \hline
    intercept & 3.68 & 0.51 & 2.68 & 3.68 & 4.69 & TRUE \\ 
      Spr & -0.07 & 0.08 & -0.23 & -0.07 & 0.10 & FALSE \\ 
      Sum & -0.55 & 0.09 & -0.72 & -0.55 & -0.38 & TRUE \\ 
      Win & -0.23 & 0.09 & -0.41 & -0.23 & -0.05 & TRUE \\ 
      BC & 0.82 & 0.16 & 0.51 & 0.82 & 1.14 & TRUE \\ 
      M & -1.92 & 0.92 & -3.75 & -1.91 & -0.11 & TRUE \\ 
      NB & -2.63 & 1.05 & -4.72 & -2.63 & -0.57 & TRUE \\ 
      NL & -3.08 & 1.11 & -5.31 & -3.06 & -0.89 & TRUE \\ 
      NT & -5.41 & 1.27 & -7.92 & -5.39 & -2.91 & TRUE \\ 
      NS & -2.42 & 1.00 & -4.39 & -2.41 & -0.46 & TRUE \\ 
      N & -5.01 & 1.27 & -7.52 & -5.01 & -2.53 & TRUE \\ 
      O & 5.61 & 2.45 & 0.80 & 5.59 & 10.51 & TRUE \\ 
      PEI & -4.13 & 1.22 & -6.54 & -4.12 & -1.73 & TRUE \\ 
      Q & 2.20 & 0.93 & 0.36 & 2.19 & 4.07 & TRUE \\ 
      S & -2.59 & 0.94 & -4.46 & -2.57 & -0.75 & TRUE \\ 
      Y & -4.15 & 1.25 & -6.61 & -4.14 & -1.71 & TRUE \\ 
      retail & -1.86 & 1.07 & -3.98 & -1.85 & 0.23 & FALSE \\ 
       \hline
    \end{tabular}
\end{table}
\subsection{Smooth Bootstrap}
The smooth bootstrap is not an ``ideal'' method for the given dataset, as only one predictor \textbf{(retail)} was continuous and real-valued. However, results were consistent with other methods, as discussed later.

Again using $B = 10000$, an algorithm was written to perform a smooth bootstrap. The \textbf{retail} column was found to have a sample variance of 1, due to the fact that it was standardized prior to model building. A reasonable value for the noise term was chosen, that is, $\frac{1}{\sqrt{n}} \approx 0.05783$. On each iteration, some $\varepsilon_i$ was added to each row $i$, where $\varepsilon \sim N(0, 0.05783)$. Using this ``new'' dataset, a model was fit and the paramter estimates were recorded. Relevant code is shown below.

\subsubsection*{Procedure}
\begin{lstlisting}
addNoise <- function(X) {

    cols <- colnames(X)
    new_X <- X
    for (col in cols) {
        Xi <- X[, col]
        if (class(data[, col]) != "factor") {
            n <- length(Xi)
            S_sq <- var(Xi)
            noise_var <- S_sq / n
            new_X[, col] <- Xi + rnorm(n=n, mean=0, sd=sqrt(noise_var))
        } else {
            new_X[, col] <- Xi
        }
    }
    return(new_X)
}

smoothBoot <- function(X, y, B, noisefunc) {

    # Get sample size
    n <- nrow(X)
    # Initialize empty vector
    params <- c()

    # Initialize progress bar
    pb <- txtProgressBar(min = 0, max = B, style = 3)

    # Perform B iterations
    for (b in 1:B) {
        # Update progress bar
        setTxtProgressBar(pb, b)

        # Get new dataset
        new_X <- noisefunc(X)
        new_data <- data.frame(protests=y, new_X)

        # Fit the model with the simulated data
        smoothboot_model <- glm.nb(protests ~., data=new_data, init.theta = 5)
        boot_params <- coef(smoothboot_model)
        params <- rbind(params, boot_params)
    }

    # Close progress bar
    close(pb)

    return(params)
}
\end{lstlisting}
\subsubsection*{Results}
\begin{table}[hbt!]
    \centering
    \begin{tabular}{|l|c|c|c|c|c|c|}
      \hline
     & mean & sd & 2.5\% & 50\% & 97.5\% & sig \\ 
      \hline
    intercept & 3.00 & 0.21 & 2.58 & 3.01 & 3.42 & TRUE \\ 
      Spr & -0.04 & 0.01 & -0.07 & -0.04 & -0.03 & TRUE \\ 
      Sum & -0.55 & 0.01 & -0.56 & -0.55 & -0.54 & TRUE \\ 
      Win & -0.20 & 0.01 & -0.23 & -0.20 & -0.18 & TRUE \\ 
      BC & 0.67 & 0.05 & 0.57 & 0.67 & 0.76 & TRUE \\ 
      M & -0.67 & 0.39 & -1.43 & -0.68 & 0.10 & FALSE \\ 
      NB & -1.21 & 0.44 & -2.08 & -1.22 & -0.33 & TRUE \\ 
      NL & -1.56 & 0.47 & -2.49 & -1.57 & -0.62 & TRUE \\ 
      NT & -3.68 & 0.52 & -4.70 & -3.68 & -2.63 & TRUE \\ 
      NS & -1.07 & 0.42 & -1.89 & -1.08 & -0.23 & TRUE \\ 
      N & -3.30 & 0.53 & -4.33 & -3.30 & -2.24 & TRUE \\ 
      O & 2.24 & 1.05 & 0.14 & 2.25 & 4.29 & TRUE \\ 
      PEI & -2.47 & 0.51 & -3.48 & -2.48 & -1.45 & TRUE \\ 
      Q & 0.92 & 0.40 & 0.13 & 0.92 & 1.69 & TRUE \\ 
      S & -1.31 & 0.40 & -2.09 & -1.32 & -0.52 & TRUE \\ 
      Y & -2.46 & 0.52 & -3.49 & -2.47 & -1.41 & TRUE \\ 
      retail & -0.39 & 0.45 & -1.29 & -0.40 & 0.52 & FALSE \\ 
       \hline
    \end{tabular}
\end{table}

\subsection{Error-Sampling Bootstrap}
Another bootstrap method was implemented, in which the error terms from the fitted model were randomly sampled with replacement, and added to the fitted values. Notably, some resulting simulated counts were rounded up to zero in the case where a negative value was produced. This was required both logically; as protests counts cannot be negative, and mathematically; as the negative binomial glm cannot be fit with negative training outputs. As a result, the integrity of the simulated datasets was not assumed to be completely intact. That being said, the results were quite consistent with the previous methods.

\subsubsection*{Procedure}

\begin{lstlisting}
epsilonBoot <- function(X, model, B, errors) {

    # Get sample size
    n <- nrow(X)
    # Initialize empty vector
    params <- c()
    # Perform B iterations
    for (b in 1:B) {
        # Get errors
        errs <- sample(errors, size=n, replace=TRUE)
        # Get fitted values
        yhat <- fitted(model)
        # Get simulated y
        ystar <- yhat + errs
        # round up negative values
        ystar <- pmax(rep(0, n), ystar)
        # Turn into DataFrame
        sim_data <- data.frame(protests=ystar, X)
        # Fit the model with the simulated data
        paramboot_model <- glm.nb(protests ~., data=sim_data, init.theta = 5)
        boot_params <- coef(paramboot_model)
        params <- rbind(params, boot_params)
    }
    return(params)
}
\end{lstlisting}
\subsubsection*{Results}
\begin{table}[hbt!]
    \centering
    \begin{tabular}{|l|c|c|c|c|c|c|}
      \hline
     & mean & sd & 2.5\% & 50\% & 97.5\% & sig \\ 
      \hline
    intercept & 3.68 & 0.51 & 2.68 & 3.68 & 4.69 & TRUE \\ 
      Spr & -0.07 & 0.08 & -0.23 & -0.07 & 0.10 & FALSE \\ 
      Sum & -0.55 & 0.09 & -0.72 & -0.55 & -0.38 & TRUE \\ 
      Win & -0.23 & 0.09 & -0.41 & -0.23 & -0.05 & TRUE \\ 
      BC & 0.82 & 0.16 & 0.51 & 0.82 & 1.14 & TRUE \\ 
      M & -1.92 & 0.92 & -3.75 & -1.91 & -0.11 & TRUE \\ 
      NB & -2.63 & 1.05 & -4.72 & -2.63 & -0.57 & TRUE \\ 
      NL & -3.08 & 1.11 & -5.31 & -3.06 & -0.89 & TRUE \\ 
      NT & -5.41 & 1.27 & -7.92 & -5.39 & -2.91 & TRUE \\ 
      NS & -2.42 & 1.00 & -4.39 & -2.41 & -0.46 & TRUE \\ 
      N & -5.01 & 1.27 & -7.52 & -5.01 & -2.53 & TRUE \\ 
      O & 5.61 & 2.45 & 0.80 & 5.59 & 10.51 & TRUE \\ 
      PEI & -4.13 & 1.22 & -6.54 & -4.12 & -1.73 & TRUE \\ 
      Q & 2.20 & 0.93 & 0.36 & 2.19 & 4.07 & TRUE \\ 
      S & -2.59 & 0.94 & -4.46 & -2.57 & -0.75 & TRUE \\ 
      Y & -4.15 & 1.25 & -6.61 & -4.14 & -1.71 & TRUE \\ 
      retail & -1.86 & 1.07 & -3.98 & -1.85 & 0.23 & FALSE \\ 
       \hline
    \end{tabular}
\end{table}

\subsection{Method Comparison}

\newpage
\section{Monte Carlo Estimation}

\newpage
\section{Conclusion}

\end{document}